Model Comparisons

### 2015-11-28 models/Model_Comparison.Rmd
Minor tweaks - prelim tests between unnormalized and normalized

# Run model functionality
```{r}
# run.partial()
# ------------
# Run RSA with model1 (entailment) and model2 (entailment + generic)
run.partial = function(d, alpha = 1, useprior = F, usenone = F, normalize=F) {
  if (normalize) {
    mat = d %>%
      select(stars, degree, speaker.p) %>%
      spread(degree, speaker.p) %>%
      mutate(hi = hi / sum(hi),
             low = low / sum(low)) %>%
      select(hi, low)
  } else {
    mat = d %>%
      select(stars, degree, speaker.p) %>%
      spread(degree, speaker.p) %>%
      select(hi, low)
  }

@@

## Tune hyperparams
```{r}
# tune.alhpa()
# ------------
# d        --> data
# alphas   --> range of alphas to test
# type     --> full or partial model
# useprior --> use uniform prior
# usenone  --> use gener None
tune.alpha = function(d, alphas = seq(from=1, to=10),
                      type="partial", useprior = T,
                      usenone=F, compare.data=NULL, addMid = F, normalize=F) {
  # Tune best alphas
  fit = sapply(alphas, FUN=function(n) {
    if (type == "partial") {
      md = d %>%
        do(run.partial(., alpha = n, useprior = useprior, usenone = usenone, normalize=normalize))
    } else {
      md = d %>%
        do(run.full(., alpha = n, useprior = useprior, usenone = usenone, addMid=addMid, normalize=normalize))
    }
    # Fit to e6 data
    if (!is.null(compare.data)) {
      matched.items = which((md[, "scale"] != "some_all" &
                               (md[, "degree"] == "hi2" | md[, "degree"] == "hi1")) |
                              (md[, "scale"] == "some_all" &
                                 (md[, "degree"] == "hi1" | md[, "degree"] == "low1")))
      md = md[matched.items, ]
      md$degree = ifelse(md$degree == "hi1", "hi", "low")
      stopifnot("listener.p" %in% colnames(compare.data))
      md$listener.p = compare.data$listener.p
    }

@@

## Data for entailment models
```{r entailment_models_data, message=FALSE, warning=FALSE}
speaker = read.csv("~/Desktop/Projects/scalar_implicature/models/model_data/L0_e8.csv")
listener = read.csv("~/Desktop/Projects/scalar_implicature/models/model_data/L1_e6.csv")
# Uniform priors
priors = data.frame(stars = seq(1, 5), prior.p = rep(0.2, 5))
# Combine speaker / listener
data.partial = left_join(speaker, listener) %>%
  left_join(priors) %>%
  rowwise %>%
  select(scale, degree, stars, speaker.p, listener.p, prior.p) %>%
  mutate(listener.p = ifelse(is.na(listener.p), 0, listener.p)) %>%
  group_by(scale)
```

## Data for full model
```{r full_model_data, message=FALSE, warning=FALSE}
speaker = read.csv("~/Desktop/Projects/scalar_implicature/models/model_data/L0_e10.csv")
listener = read.csv("~/Desktop/Projects/scalar_implicature/models/model_data/L1_e11.csv")
data.full = left_join(speaker, listener) %>%
  left_join(priors) %>%
  rowwise %>%
  select(scale, degree, stars, speaker.p, listener.p, prior.p) %>%
  mutate(listener.p = ifelse(is.na(listener.p), 0, listener.p)) %>%
  group_by(scale)

speaker = read.csv("~/Desktop/Projects/scalar_implicature/models/model_data/L0_e10a.csv")
data.full.extras = left_join(speaker, listener) %>%
  left_join(priors) %>%
  rowwise %>%
  select(scale, degree, stars, speaker.p, listener.p, prior.p) %>%
  mutate(listener.p = ifelse(is.na(listener.p), 0, listener.p)) %>%
  group_by(scale)
```

@@

## Run comparisons
```{r run_comparisons, message=FALSE}
# Save performance output
performance.output = data.frame(model=rep(NA, 8), cor=rep(NA, 8))
alphas = rep(NA, 8)

# Model 1 - Entailment, no alpha tuning
# -------------------------------------
alphas[1] = 1
m1.noFit = data.partial %>%
  do(run.partial(., alpha = alphas[1], useprior = T, usenone = F))
performance.output[1, ] = c("M1_noFit", round(cor(m1.noFit$listener.p, m1.noFit$pred), 5))

# Model 1 - Entailment, alpha tuning
# ----------------------------------
alphas[2] = tune.alpha(data.partial)
m1.fit = data.partial %>%
  do(run.partial(., alpha = alphas[2]))
performance.output[2, ] = c("M1_fit", round(cor(m1.fit$listener.p, m1.fit$pred), 5))

@@

# Model 4 - Full, no alpha tuning
# -------------------------------
alphas[7] = 1
m4.noFit = data.full.extras %>%
  do(run.full(., alpha = alphas[7], addMid=T))
m4.noFit.matched = m4.noFit[matched.items.extras, ]
# Need to fix this to dealt with 'mid'
m4.noFit.matched$degree = ifelse(m4.noFit.matched$degree == "hi1", "hi", "low")
m4.noFit.matched = cbind(m4.noFit.matched, data.partial$listener.p)
# Check that everything lines up between partial and full
all(m4.noFit.matched$scale == data.partial$scale & m4.noFit.matched$degree == data.partial$degree)
colnames(m4.noFit.matched)[length(colnames(m4.noFit.matched))] = "e6.listener.p"
# store corr
performance.output[7, ] = c("M4_noFit", round(cor(m4.noFit.matched$e6.listener.p, m4.noFit.matched$pred), 5))

>>>>

# Run model functionality
```{r}
# run.partial()
# ------------
# Run RSA with model1 (entailment) and model2 (entailment + generic)
run.partial = function(d, alpha=1, useprior=F, usenone=F, normalize=F) {
  if (normalize) {
    mat = d %>%
      select(stars, degree, speaker.p) %>%
      spread(degree, speaker.p) %>%
      mutate(hi = hi / sum(hi),
             low = low / sum(low)) %>%
      select(hi, low)
  } else {
    mat = d %>%
      select(stars, degree, speaker.p) %>%
      spread(degree, speaker.p) %>%
      select(hi, low)
  }

@@

## Tune hyperparams
```{r}
# tune.alhpa()
# ------------
# d        --> data
# alphas   --> range of alphas to test
# type     --> full or partial model
# useprior --> use uniform prior
# usenone  --> use gener None
tune.alpha = function(d, alphas = seq(from=1, to=10),
                      type="partial", useprior = T,
                      usenone=F, compare.data=NULL, addMid = F, normalize=F) {
  # Tune best alphas
  fit = sapply(alphas, FUN=function(n) {
    if (type == "partial") {
      md = d %>%
        do(run.partial(., alpha=n, useprior=useprior, usenone=usenone, normalize=normalize))
    } else {
      md = d %>%
        do(run.full(., alpha=n, useprior=useprior, usenone=usenone, addMid=addMid, normalize=normalize))
    }
    # Fit to e6 data
    if (!is.null(compare.data)) {
      matched.items = which((md[, "scale"] != "some_all" &
                               (md[, "degree"] == "hi2" | md[, "degree"] == "hi1")) |
                              (md[, "scale"] == "some_all" &
                                 (md[, "degree"] == "hi1" | md[, "degree"] == "low1")))
      md = md[matched.items, ]
      md$degree = ifelse(md$degree == "hi1", "hi", "low")
      stopifnot("listener.p" %in% colnames(compare.data))
      md$listener.p = compare.data$listener.p
    }

@@

### Data for entailment models
```{r entailment_models_data, message=FALSE, warning=FALSE}
speaker = read.csv("~/Desktop/Projects/scalar_implicature/models/model_data/L0_e8.csv")
listener = read.csv("~/Desktop/Projects/scalar_implicature/models/model_data/L1_e6.csv")
# Uniform priors
priors = data.frame(stars = seq(1, 5), prior.p = rep(0.2, 5))
# Combine speaker / listener
data.partial = left_join(speaker, listener) %>%
  left_join(priors) %>%
  rowwise %>%
  select(scale, degree, stars, speaker.p, listener.p, prior.p) %>%
  mutate(listener.p = ifelse(is.na(listener.p), 0, listener.p)) %>%
  group_by(scale)
```

### Data for full model - no alternatives
```{r full_model_data, message=FALSE, warning=FALSE}
speaker = read.csv("~/Desktop/Projects/scalar_implicature/models/model_data/L0_e10.csv")
listener = read.csv("~/Desktop/Projects/scalar_implicature/models/model_data/L1_e11.csv")
data.full = left_join(speaker, listener) %>%
  left_join(priors) %>%
  rowwise %>%
  select(scale, degree, stars, speaker.p, listener.p, prior.p) %>%
  mutate(listener.p = ifelse(is.na(listener.p), 0, listener.p)) %>%
  group_by(scale)
```

### Data for full model - alternatives
```{r}
speaker = read.csv("~/Desktop/Projects/scalar_implicature/models/model_data/L0_e10a.csv")
data.full.extras = left_join(speaker, listener) %>%
  left_join(priors) %>%
  rowwise %>%
  select(scale, degree, stars, speaker.p, listener.p, prior.p) %>%
  mutate(listener.p = ifelse(is.na(listener.p), 0, listener.p)) %>%
  group_by(scale)
```

@@

## Run comparisons
```{r run_comparisons, message=FALSE}
# Save performance output
performance.output = data.frame(model=rep(NA, 8), norm.cor=rep(NA, 8))
alphas = rep(NA, 8)

# Model 1 - Entailment only model
# --------------------------------
# a)
# Normalized = T
# alpha tune = F
alphas[1] = 1
m1.a = data.partial %>%
  do(run.partial(., alpha = alphas[1], useprior=T, usenone=F, normalize=F))
performance.output[1, ] = c("M1_noFit", round(cor(m1.noFit$listener.p, m1.noFit$pred), 5))
# b)
# Normalized = F
# alpha tune = F
alphas[1] = 1
m1.b = data.partial %>%
  do(run.partial(., alpha = alphas[1], useprior=T, usenone=F, normalize=F))
performance.output[1, ] = c("M1_noFit", round(cor(m1.noFit$listener.p, m1.noFit$pred), 5))

# Model 1 - Entailment, alpha tuning
# ----------------------------------
alphas[2] = tune.alpha(data.partial, normalize=F)
m1.fit = data.partial %>%
  do(run.partial(., alpha = alphas[2]))
performance.output[2, ] = c("M1_fit", round(cor(m1.fit$listener.p, m1.fit$pred), 5))

@@

# Model 4 - Full, no alpha tuning
# -------------------------------
alphas[7] = 1
m4.noFit = data.full.extras %>%
  do(run.full(., alpha = alphas[7], addMid=T, normalize=T))
m4.noFit.matched = m4.noFit[matched.items.extras, ]
# Need to fix this to dealt with 'mid'
m4.noFit.matched$degree = ifelse(m4.noFit.matched$degree == "hi1", "hi", "low")
m4.noFit.matched = cbind(m4.noFit.matched, data.partial$listener.p)
# Check that everything lines up between partial and full
all(m4.noFit.matched$scale == data.partial$scale & m4.noFit.matched$degree == data.partial$degree)
colnames(m4.noFit.matched)[length(colnames(m4.noFit.matched))] = "e6.listener.p"
# store corr
performance.output[7, ] = c("M4_noFit", round(cor(m4.noFit.matched$e6.listener.p, m4.noFit.matched$pred), 5))
